Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-08-02 doc/manuscript/manuscript.Rmd
update versions, auto generate poppr table

All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in R version 3.4.0 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
<!-- add something about repeat length consistency  -->
Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To avoid including samples that were potentially collected from the same plant (which increases the probability of collecting the same sample more than once), data were clone-corrected on the region/source/host/year hierarchy for subsequent analyses. The clone-corrected data set included 318 isolates.

@@

We wanted to assess if there was correlation between multilocus genotypes (MLG) and mycelial compatibility groups (MCG). 
This was done using a graph-based approach with community analysis.
We first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection. 
This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLGs.
The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.0.1 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
All communities that had fewer than 10 members were then consolidated into one.
These community definitions were then used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].

@@

----------------------------------------------------------------------------
Pop             N        eMLG    H    G  $\lambda$  $E_5$   *h* $I_A$ $\bar{ 
                                                                       r}_d$ 
--------- ------- ----------- ---- ---- ---------- ------ ----- ----- ------
WA        58 (56) 9.95 (0.23)  4.0 54.3       0.98   0.98  0.60 0.65*  0.07*

MI        58 (43)  9.3 (0.79)  3.6 29.0       0.97   0.78  0.54 1.25*  0.14*

ND        41 (35) 9.44 (0.73)  3.5 25.9       0.96   0.82  0.54    1*   0.1*

NE        37 (28) 8.93 (0.94)  3.2 17.8       0.94   0.75  0.55 2.22*  0.25*

CO        34 (28) 9.46 (0.67)  3.3 24.1       0.96   0.92  0.56 2.33*  0.27*

France    21 (14)  8.5 (0.85)  2.6 12.6       0.92   0.95  0.48 0.94*  0.11*

CA        18 (15) 9.12 (0.72)  2.7 13.5       0.93   0.94  0.51 0.28~  0.03~

OR        17 (13) 8.52 (0.85)  2.5 10.7       0.91   0.89  0.47 0.92*   0.1*

Mexico     15 (9)  7.1 (0.85)  2.1  7.3       0.86   0.89  0.28 2.57*  0.37*

MN          9 (7)       7 (0)  1.9  6.2       0.84   0.93  0.47 1.34*  0.19*

Australia   6 (6)       6 (0)  1.8  6.0       0.83   1.00  0.48  0.7~  0.12~

WI          2 (2)       2 (0)  0.7  2.0       0.50   1.00  0.27     -      -

NY          1 (1)       1 (0)  0.0  1.0       0.00    NaN   NaN     -      -

ID          1 (1)       1 (0)  0.0  1.0       0.00    NaN   NaN     -      -

----------------------------------------------------------------------------

Table: (\#tab:iatab) Genotypic diversity and Linkage Disequilibrium summary for geographic populations arranged by abundance after clone-correction by a hierarchy of Region/Source/Host/Year. 
Pop = Population, N = number of individuals (number of MLG in parentheses), 
eMLG = expected number of MLGs based on rarefaction at min(10, N) individuals (standard error in parentheses), 
H = Shannon-Weiner Index, G = Stoddardt and Taylor's Index, 
$\lambda$ = Simpson's Index, 
*h* = Nei's 1978 gene diversity, 
$E_5$ = Evenness,
$I_A$ = the index of association, 
$\bar{r}_d$ = standardized index of association. 
P-value symbols: \* = 0.001, ~ = $\leq 0.05$.

>>>>

All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in R version 3.4.1 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
<!-- add something about repeat length consistency  -->
Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To avoid including samples that were potentially collected from the same plant (which increases the probability of collecting the same sample more than once), data were clone-corrected on the region/source/host/year hierarchy for subsequent analyses. The clone-corrected data set included 318 isolates.

@@

We wanted to assess if there was correlation between multilocus genotypes (MLG) and mycelial compatibility groups (MCG). 
This was done using a graph-based approach with community analysis.
We first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection. 
This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLGs.
The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.1.2 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
All communities that had fewer than 10 members were then consolidated into one.
These community definitions were then used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].

@@

```{r table_ia, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
suppressPackageStartupMessages(library("tidyverse"))
read_csv("../../results/tables/genotype_diversity_table_region.csv", 
         na = "NA", # spaces read in as spaces
         col_types = cols(
         Pop = col_character(),
         N = col_integer(),
         MLG = col_integer(),
         eMLG = col_double(),
         SE = col_double(),
         H = col_double(),
         G = col_double(),
         lambda = col_double(),
         E.5 = col_double(),
         Hexp = col_double(),
         Ia = col_double(),
         p.Ia = col_double(),
         rbarD = col_double(),
         p.rD = col_double(),
         File = col_character()
      )) %>%
  mutate(Ia = case_when(
    !is.finite(p.Ia) ~ "-",
    p.Ia == 0.001 ~ paste0(round(Ia, 2), "*"),
    p.Ia <= 0.01  ~ paste0(round(Ia, 2), "-"),
    p.Ia <= 0.05  ~ paste0(round(Ia, 2), "~"),
    TRUE          ~ as.character(round(Ia, 2))
  )) %>%
  mutate(rbarD = case_when(
    !is.finite(p.rD) ~ "-",
    p.rD == 0.001 ~ paste0(round(rbarD, 2), "*"),
    p.rD <= 0.01  ~ paste0(round(rbarD, 2), "-"),
    p.rD <= 0.05  ~ paste0(round(rbarD, 2), "~"),
    TRUE          ~ as.character(round(rbarD, 2))
  )) %>%
  mutate_if(is.numeric, round, digits = 2) %>%
  mutate(N = paste(N, MLG, sep = " (") %>% paste0(")") %>% formatC()) %>%
  mutate(eMLG = paste(eMLG, SE, sep = " (") %>%  paste0(")") %>% formatC()) %>%
  rename(`*h*` = Hexp) %>%
  rename(`$E_5$` = E.5) %>%
  rename(`$I_A$` = Ia) %>%
  rename(`$\\lambda$` = lambda) %>%
  rename(`$\\bar{r}_d$` = rbarD) %>%
  select(-File, -MLG, -SE, -p.rD, -p.Ia) %>% 
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(huxtable::everywhere, 4:5, 1) %>%
  huxtable::set_align(huxtable::everywhere, -1, "right") %>%
  huxtable::set_col_width(c(1.1, 0.9, 1.3, 0.5, 0.5, 1.2, 0.7, 0.6, 0.6, 0.8)) %>%
  huxtable::print_md(max_width = 80)
```

Table: (\#tab:iatab) Genotypic diversity and Linkage Disequilibrium summary for geographic populations arranged by abundance after clone-correction by a hierarchy of Region/Source/Host/Year. 
Pop = Population, N = number of individuals (number of MLG in parentheses), 
eMLG = expected number of MLGs based on rarefaction at min(10, N) individuals (standard error in parentheses), 
H = Shannon-Weiner Index, G = Stoddardt and Taylor's Index, 
$\lambda$ = Simpson's Index, 
*h* = Nei's 1978 gene diversity, 
$E_5$ = Evenness,
$I_A$ = the index of association, 
$\bar{r}_d$ = standardized index of association. 
P-value symbols: \* = 0.001, - = $\leq 0.01$, ~ = $\leq 0.05$.
